---
title: "HLM"
author: "Dr. Immanuel Williams"
date: "July 17, 2019"
output: html_document
---



```{r p0_lib_data, message=FALSE, warning=FALSE, include=FALSE}
## Library
library(tidyverse)   ## Data Manipulation
library(lme4)        ## Multilevel Model
library(lmerTest)    ## Find p-values for lmer models
library(kableExtra)

## Data
setwd('C:/Users/james/OneDrive/Documents/Important_Files/Old Projects/UTQ_Battey/Data Analysis/Attempt 25(new paper)/all_teachers_analysis')
df = read.csv('sciach.csv')
df$GROUP = as.factor(df$GROUP)
```




```{r p0_group_means, echo=FALSE, message=FALSE, warning=FALSE}
GROUP_means = df %>%
group_by(GROUP) %>%
summarise(mean_group = mean(SCIENCE, na.rm = T)) %>%
select(mean_group)

GROUP_means %>%
  kable() %>%
  kable_styling(position = "center", full_width = F,bootstrap_options ="bordered")
```


### Model 1: Analysis of Variance Model (ANOVA)
ANOVA (*model_1*):
$$y_{ij} = \mu + \alpha_j + \epsilon_{ij} $$

where $j = 1, ..., k$, $i = 1, ..., n_j$ and $N(0,\sigma^2)$

- $y_{ij}$: response variable
- $\mu$: Overall Mean
- $\alpha_j$: $j$th mean
- $\epsilon_{ij}$: error

### Model 2: Multilevel Model - Random Intercepts
The ANOVA Model can be be rexpressed as a Multilevel ANOVA. (*model_2*):

This expression can be seen as a **Level-1** model:
$$y_{ij} = \beta_{0j} + e_{ij} $$

and **Level-2** model:
$$\beta_{0j} = \gamma_{00} + u_{0j}$$

All together:
$$y_{ij} = \gamma_{00} + u_{0j} + e_{ij} $$


- $y_{ij}$: response variable
- $\gamma_{00}$:  $y$-intercept or the expected value when the covariate is equal 0
- $u_{0j}$: Level-2 error
- $e_{ij}$: Level-1 error


```{r p1_model_1, echo=FALSE, message=FALSE, warning=FALSE}
model_1 = aov(SCIENCE ~ GROUP, df)
summary(model_1)
```




```{r p1_model_2, echo=FALSE, message=FALSE, warning=FALSE}
model_2 = lmer(SCIENCE ~ (1|GROUP),data = df)
summary(model_2) 
```






```{r p1_means_comp_1, echo=FALSE, message=FALSE, warning=FALSE}
## Create df of means
means_model2 = data.frame(true_means = GROUP_means,
                     ran_int_means = coef(model_2)$GROUP) 

colnames(means_model2) = c('True_Means','Ran_Int_Means')
## View Diff
means_model2 %>%
mutate(Diff = round(True_Means - Ran_Int_Means,3))
```






```{r view_variance_model2, echo=FALSE, message=FALSE, warning=FALSE}
boxplot(resid(model_2) ~ GROUP, df, horizontal = T,
        main = "Homogeneity of Variance Model 2")
```



### Model 3: Regression Model (*model_3*)

$$y_{ij} = \beta_0 + \beta_1 x_{ij} + e_{ij} $$

- $y_{ij}$: response variable
- $\beta_0$: y-intercept coefficient
- $\beta_1$: slope coefficient
- $x_{ij}$: explanatory variable
- $e_{ij}$: Level-1 error

### Model 4:  Random Intercepts with Covariate (*model_4*)

$$y_{ij} = \gamma_{00} + \gamma_{10} x_{ij} + u_{0j} + e_{ij} $$

- $y_{ij}$: response variable
- $\gamma_{00}$:  $y$-intercept or the expected value when the covariate is equal 0
- $\gamma_{10}$: he expected change in the response variable ($y$) for every one unit change in the covariate
- $x_{ij}$: explanatory variable
- $u_{0j}$: Level-2 error
- $e_{ij}$: Level-1 error

```{r p2_model_3, echo=FALSE, message=FALSE, warning=FALSE}
model_3 =   lm(SCIENCE ~ URBAN, df)
summary(model_3)
```




```{r p2_model_4, echo=FALSE, message=FALSE, warning=FALSE}
model_4 = lmer(SCIENCE ~ URBAN + (1|GROUP),data = df)
summary(model_4) 
```




## Comparison Random Intercept vs. Random Intercepts with a Covariate

```{r p2_model_fit_comp_2, echo=FALSE, message=FALSE, warning=FALSE}
anova(model_2,model_4)
```



### Model 5:  Random Intercepts  and Slope with Covariate (*model_5*)

$$y_{ij} = \gamma_{00} + \gamma_{10} x_{ij} + u_{0j} + u_{1j} + e_{ij} $$

- $y_{ij}$: response variable
- $\gamma_{00}$:  $y$-intercept or the expected value when the covariate is equal 0
- $\gamma_{10}$: he expected change in the response variable ($y$) for every one unit change in the covariate
- $x_{ij}$: explanatory variable
- $u_{0j}$: Level-2 error
- $u_{1j}$: Level-2 error Covariate
- $e_{ij}$: Level-1 error



### Model 6:  Random Intercepts and Slope with Covariate (*model_6*)

$$y_{ij} = \gamma_{00} + \gamma_{10} x_{ij} +\gamma_{01} x_{.j} + u_{0j} + u_{1j} + e_{ij} $$

- $y_{ij}$: response variable
- $\gamma_{00}$:  $y$-intercept or the expected value when the covariate is equal 0
- $\gamma_{10}$: he expected change in the response variable ($y$) for every one unit change in the covariate
- $\gamma_{01}$: 
- $x_{.j}$:
- $x_{ij}$: explanatory variable
- $u_{0j}$: Level-2 error
- $u_{1j}$: Level-2 error Covariate
- $e_{ij}$: Level-1 error



```{r p3_model_5, echo=FALSE, message=FALSE, warning=FALSE}
model_5 <- lmer(SCIENCE ~ URBAN + (URBAN | GROUP),data = df)
summary(model_5)
```




```{r p3_model_6, echo=FALSE, message=FALSE, warning=FALSE}
model_6 = lmer(SCIENCE ~ URBAN + SCH.RES + (URBAN |GROUP), data = df)
## URBAN - x; SCIENCE - y; SCH.RES - Group x; GROUP - ID
summary(model_6) 
```



## Comparison of All models
```{r p3_model_fit_comp_3, echo=FALSE, message=FALSE, warning=FALSE}
anova(model_2,model_4,model_5,model_6)
```


























