---
title: "HLM"
author: "Dr. Immanuel Williams"
date: "July 17, 2019"
output: html_document
---



```{r p0_lib_data, message=FALSE, warning=FALSE, include=FALSE}
## Library
library(tidyverse)   ## Data Manipulation
library(lme4)        ## Multilevel Model
library(lmerTest)    ## Find p-values for lmer models
library(kableExtra)

## Data
setwd('C:/Users/james/OneDrive/Documents/Important_Files/Old Projects/UTQ_Battey/Data Analysis/Attempt 25(new paper)/all_teachers_analysis')
df = read.csv('sciach.csv')
df$GROUP = as.factor(df$GROUP)
```




```{r p0_group_means, echo=FALSE, message=FALSE, warning=FALSE}
GROUP_means = df %>%
group_by(GROUP) %>%
summarise(mean_group = mean(SCIENCE, na.rm = T)) %>%
select(mean_group)

GROUP_means %>%
  kable() %>%
  kable_styling(position = "center", full_width = F,bootstrap_options ="bordered")
```


### Model 1: Analysis of Variance Model (ANOVA)
ANOVA (*model_1*):
$$y_{ij} = \mu + \alpha_j + \epsilon_{ij} $$

where $j = 1, ..., k$, $i = 1, ..., n_j$ and $N(0,\sigma^2)$

- $y_{ij}$: response variable
- $\mu$: Overall Mean
- $\alpha_j$: $j$th mean
- $\epsilon_{ij}$: error

### Model 2: Multilevel Model - Random Intercepts
The ANOVA Model can be be rexpressed as a Multilevel ANOVA. (*model_2*):

This expression can be seen as a **Level-1** model:
$$y_{ij} = \beta_{0j} + e_{ij} $$

and **Level-2** model:
$$\beta_{0j} = \gamma_{00} + u_{0j}$$

All together:
$$y_{ij} = \gamma_{00} + u_{0j} + e_{ij} $$


- $y_{ij}$: response variable
- $\gamma_{00}$:  $y$-intercept or the expected value when the covariate is equal 0
- $u_{0j}$: Level-2 error
- $e_{ij}$: Level-1 error


```{r p1_model_0, echo=FALSE, message=FALSE, warning=FALSE}
model_0 = aov(SCIENCE ~ GROUP, df)
summary(model_0)
```

